---
title: "Movielens-Project-Report"
author: "Harshad B."
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include=FALSE}
options(tinytex.verbose = TRUE)
library(knitr)
library(caret)
library(tidyverse)
library(data.table)
edx <- readRDS("D:/r-projects/edx.rds")
validation <- readRDS("D:/r-projects/validation.rds")
```

## Introduction
Machine learning helps us to describe data and deduce useful patterns from the same. The aim of machine learning is to process data into helpful information and naturally intuitive solutions. In 2006, Netflix placed a seven-figure bounty on a verified improvemnent to their movie recommendation system.

The following project is based on the Netflix Challenge. In this project, we use the [MovieLens 10M dataset](https://grouplens.org/datasets/movielens/10m/) that consists of 10 million ratings and 100,000 tag applications applied to 10,000 movies by 72,000 users.

The dataset is divided into edx and validation sets in a 90-10 ratio.

## Approach:

First, the edx set has been divided into two sets: edx_train and edx_test. Various models will be created using the _edx_train_ set and their RMSEs will be calculated on the _edx_test_ set. When a model with RMSE close to the expected RMSE is achieved, then the _edx_ set will be used to train and predict on the _vaildation_ set.

### Partitioning of _edx_ dataset
``` {r data-partition, echo = TRUE}
testindex <- createDataPartition(edx$rating, times = 1, p = 0.2, list = FALSE)
edx_train <- edx[-testindex,]
edx_test <- edx[testindex,]
edx_test <- edx_test %>%
  semi_join(edx_train, by = "movieId") %>%
  semi_join(edx_train, by = "userId")
```

## Method :
### Calculating rmse of Average Method on _edx_test_ set
``` {r avg-method, echo = TRUE}
mean_tt <- mean(edx_train$rating)
rmse_avg <- RMSE(edx_test$rating, mean_tt)
```

## Results
``` {r avg-rmse, echo = FALSE}
rmse_results <- tibble(Method = "Just the Average (edx_test)", RMSE = rmse_avg)
rmse_results %>% kable()
```

## Method :
### Calculating rmse of movie effect on _edx_test_ set
``` {r movie-method, echo = TRUE}
movie_tt <- edx_train %>%
  group_by(movieId) %>%
  summarize(bi = mean(rating - mean_tt), .groups = 'drop')
pred_bi <- mean_tt + edx_test %>%
  left_join(movie_tt, by='movieId') %>%
  .$bi
rmse_movie <- RMSE(pred_bi, edx_test$rating)
```

## Results
``` {r movie-rmse, echo = FALSE}
rmse_results <- bind_rows(rmse_results, tibble(Method="Movie Effect Model (edx_test)", RMSE = rmse_movie ))
rmse_results %>% kable()
```

## Method :
### Calculating rmse of movie and user model on _edx_test_ set
``` {r movie-user-method, echo = TRUE}
user_tt <- edx_test %>%
  left_join(movie_tt, by='movieId') %>%
  group_by(userId) %>%
  summarize(bu = mean(rating - mean_tt), .groups = 'drop')
pred_bu <- edx_test %>%
  left_join(movie_tt, by='movieId') %>%
  left_join(user_tt, by='userId') %>%
  mutate(pred = mean_tt + bi + bu) %>%
  .$pred
rmse_user <- RMSE(pred_bu, edx_test$rating)
```

## Results
``` {r movie-user-rmse, echo = FALSE}
rmse_results <- bind_rows(rmse_results, tibble(Method="User + Movie Effect Model (edx_test)", RMSE = rmse_user ))
rmse_results %>% kable()
```

## Method :
### Calculating rmse of regularized movie and user model on _edx_test_ set
``` {r reg-movie-user-method, echo = TRUE}
lambdas <- seq(0, 10, 0.25)
rmses <- sapply(lambdas, function(l){
  mu <- mean(edx_train$rating)
  bi <- edx_train %>%
    group_by(movieId) %>%
    summarize(bi = sum(rating - mean_tt)/(n()+l), .groups = 'drop')
  bu <- edx_train %>%
    left_join(bi, by="movieId") %>%
    group_by(userId) %>%
    summarize(bu = sum(rating - bi - mean_tt)/(n()+l), .groups = 'drop')
  pred <-
    edx_test %>%
    left_join(bi, by = "movieId") %>%
    left_join(bu, by = "userId") %>%
    mutate(pred = mean_tt + bi + bu) %>%
    .$pred
  return(RMSE(pred, edx_test$rating))
})
```

The plot below shows us qq-plot of _lambdas vs. rmses_ for the regularized movie and user model on **edx_test** set
```{r, echo = FALSE}
qplot(lambdas, rmses)
lambdas[which.min(rmses)]
```

## Results
``` {r reg-movie-user-rmse, echo = FALSE}
rmse_results <- bind_rows(rmse_results, tibble(Method="Regularized Movie + User Effect Model (edx_test)", RMSE = min(rmses)))
rmse_results %>% kable()
```

## Method
### calculating rmse of regularized movie and user model on _validation_ set
``` {r final-method, echo = TRUE}
val_lambdas <- seq(0, 10, 0.25)
val_rmses <- sapply(lambdas, function(l){
  mu <- mean(edx$rating)
  bi <- edx %>%
    group_by(movieId) %>%
    summarize(bi = sum(rating - mu)/(n()+l), .groups = 'drop')
  bu <- edx %>%
    left_join(bi, by="movieId") %>%
    group_by(userId) %>%
    summarize(bu = sum(rating - bi - mu)/(n()+l), .groups = 'drop')
  pred <- validation %>%
    left_join(bi, by = "movieId") %>%
    left_join(bu, by = "userId") %>%
    mutate(pred = mu + bi + bu) %>%
    .$pred
  return(RMSE(pred, validation$rating))
})
final_rmse <- min(val_rmses)
```

The plot below shows us qq-plot of _val_lambdas vs. val_rmses_ for the regularized movie and user model on __validation__ set

```{r, echo = FALSE}
qplot(val_lambdas, val_rmses)
val_lambdas[which.min(rmses)]
```

## Results
``` {r final-rmse, echo = FALSE}
rmse_results <- bind_rows(rmse_results, tibble(Method="Regularized Movie + User Effect Model (validation set)", RMSE = final_rmse))
rmse_results %>% kable()
```

## Conclusion
From the above table, we can see incremental improvements to the RMSE as we supplant our model with bias terms and regularization. Because of the simplicity of the linear model, we are able to predict movie ratings without a serious toll on the computer resources.